---
title: "[wip] the age of tidyverse"
description: |
  what we can learn from music and literature about social politics in open science communities
author:
  - name: Charles T. Gray
    url: https://twitter.com/cantabile
date: 08-23-2019
bibliography: biblio.bib
output:
  distill::distill_article:
    self_contained: false
    toc: true
categories:
  - grrr...
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

[work in progress]

We like to think we're pretty enlightened and *inclusive*[@gray_coding_2019]. We've got it all worked out. We're *woke*. But a cursory glance through the literature of social politics and barriers of class inequality demonstrates humans tend to recreate the same social patterns of Othering and exclusion. 


```{r echo=TRUE,eval=FALSE}
rep(Othering, n) =
  rep(demeaning, n) =
  rep(dehumanising, n) = 
  n * trauma =
  exile
# Disadvantage creates a cascade of traumatic Othering.

```


Negative experiences in academia are far from new, consider Raskalnikov, of Dostoyevsky's 1866 *Crime and Punishment*[@dostoyevsky_crime_2004], who was driven to the titular crime by his impoverishment as an early-career academic. 

> So poorly dressed was he that another man, even one inured to such a style of living, would have been ashamed to go out in the street during the daytime in such rags. (*Crime and Punishment*[@dostoyevsky_crime_2004])


It is now standard for open science communities, large and small, to advertise a Code of Conduct, a catalogue of behaviours we consider to be unacceptable. However, we must question to what extent CoCs innoculate us from our implicit biases and perceptions of disadvantage[@gray_coding_2019]. Through film, music, and literature we consider how humans shape their world view through their lived experience, and struggle to know how to respond when culture, society, and systems that are inclusive for some fail to be inclusive for all. In particular, we consider when a culture is highly codified and structured, as it is in academia and in late nineteenth-century New York high society, and how these systems can be weaponised against those perceived as Other.

## on contemplating inequality

In Winterson's *Art & Lies*[@winterson_art_1996], the surgeon, Handel, has a comfortable life, but finds contemplating of others' struggles traumatic.

> Reportage is violence. Violence to the spirit. Violence to the emotional sympathy that should quicken in you and me when we meet face to face with pain. How many defeated among our own do we step over and push aside on our way home to watch the evening news? 'Terrible' you said at Somalia, Bosnia, Ethiopia, Russia, China, the Indian earthquake, the American floods, and then you watched a quiz show or a film because there's nothing you can do, nothing you can do, and the fear and unease that such powerlessness brings, trains in its wash, a dead arrogance for the beggar on the bridge that you pass  every day. Hasn't he got legs and a cardboard box to sleep in?

> And still we long to feel. (*Art & Lies*[@winterson_art_1996])

We can learn a lot about how our dearly held cultural values are filtered through our lived experience via an examination of social class and the barriers of inequality in music and literature.

Here we discuss three means by which minorities have been Othered, race, gender, and class, to demonstrate the mechanisms of Othering are manifold, before considering the codified world of programming communities. We first define the concept of Other, then consider  how this has played out through Othering cultures perceived as primitive, how women are Othered by being cast as irrational, and finally, we meet *The Age of Tidyverse*, where we examine on the patterns of class and inequality explored in Wharton's *The Age of Innocence* are reflected in scientific communities.

## the Other

There are many ways we can Other people. For example, Edward Said's seminal postcolonial 1979 text *Orientalism*[@trove.nla.gov.au/work/7562687] examines how western hegemony has shaped a view of those perceived as Asian, Oriental, and Other. One way this is reflected in music is a fixation on [pentatonicisation](https://en.wikipedia.org/wiki/Pentatonic_scale#Pervasiveness), a 5-note simulcrum that has come to aurally indicate Orientalism. 

![[The minor pentatonic scale](https://en.wikipedia.org/wiki/Pentatonic_scale#Pervasiveness), image: Wikipedia.](../../figs/pentatonic.png)

This musical convention has come to represent, at times interchangeably, Japanese, Chinese, Turkish, and any number of cultures perceived as *Oriental*. This interchangeability of cultures is, in itself, Othering. There is what is known, and everything else is grouped together as Other. This exacerbates bias, as we fail to consider differences between cultures that are not our own. 

The song 'Honour to us all', from Disney's 1998 animated musical *Mulan*, employs pentatonicisation to signify Orientalism. You have to admit, it sounds *Asian*, doesn't it? But does it really sound like Chinese music? Would **you** recognise the sound of an *erhu*, a ubiquitous instrument in traditional Chinese music? So, **who** decided what *sounds* Asian? 

<iframe width="560" height="315" src="https://www.youtube.com/embed/ZnUEDaeoF0Q" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

 Similarly, Saint-Sa&euml;ns' 1875 opera *Samson et Delilah*[@nla.cat-vn300872] utlises a *hijaz* scale to indicate Arabicness in a similar fashion. Puccini employed pentatonicisation in the operas *Madame Butterfly*[@nla.cat-vn1194414] (1904) and *Turandot*[@nla.cat-vn2720507] (1926), the same musical conventions employed to indicate both Japan and China. 
 
 Edith Wharton scathingly points to cultural ambiguity in opera in her melodrama *The Age of Innocence*[@wharton_age_1920](AoI), set in late-nineteenth century New York high society.
 
 > She sang, of course, "*M'ama!*", not "he loves me," since an unalterable and unquestioned law of the musical world required that the German text of French operas sung by Swedish artists should be translated into Italian for the clearer understanding of English-speaking audiences. (AoI[@wharton_age_1920])
 

 Exoticism, fascination with the Orient, swept western music throughout the nineteenth century[@locke_exoticism_2012]. Arguably the most important moment for Western exoticism was London's Great Exhibition in 1851, where Western composers were exposed to Balinese Gamelan and music from other cultures. This was an era with no sound recording, and sheet music was devilishly difficult to access. For example, Australian music history tracks ten to twenty years after European musical development, because it took that long for the sheet music to reach people[@trove.nla.gov.au/work/16869421]. Thus, the musical conventions of nineteenth century exoticism were disseminated through reproduction, diluting, at each iteration, the link to the original culture imitated.


We therefore conclude this music tells us little about the source cultures. But it speaks *volumes*, in literature literally, about how Westerner's perceive other cultures. There are many mechanisms of Othering, here we will focus on primitivism and race, and casting women as irrational as a tool of disempowerment, before examining how Othering can play out in open science communities.
 
## race: the noble savage

Within exoticism was a fascination with primitivism, reflecting the perceived as underdeveloped nature of other cultures in art. Igor Stravinsky's primitivist *The Rite of Spring* caused a riot when first performed in Paris in 1913, the "barbaric rhythms and dissonances" proving too much for the audience[@chua_rioting_2007].

<iframe width="560" height="315" src="https://www.youtube.com/embed/EkwqPJZe8ms?start=29" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

A challenge in open science is, via recommended[@bryan_happy_2019; @doi:10.1080/00031305.2017.1399928] code-sharing tools such as [GitHub](https://github.com/), the world is suddenly a much smaller place. We encounter people from other cultures far more readily than we once might have. Thus, cultural studies in text, film, and music, have arguably more importance than ever. 

Furthermore, in addition to different cultures, open science brings together researchers from disparate disciplines. In this space it is possible for a mathematician and an archaeologist to find common ground in scientific computing[@gray_truth_2019]. 

Discrimination based on race is prohibited by standard CoCs, but by considering the trope of the noble savage, we find another connection. The assumption of primitivism, and this is particularly important for researcher developers, who may well be expert researchers whose education was not in software development. 

Even whilst admiring another culture, we can bring dehumanising assumptions. In Herman Melville's 1851 epic *Moby Dick*, the protagonist Ishmael shares his room with a "peddler of heads" with a "tomahawk".

> But there was no time for shuddering, for now the savage went about something that completely fascinated my attention and conveinced me that he must be a heathen. 

Although the p


## gender: the mad woman in the attic

## class: the age of tidyverse

