---
title: "Structuring Data" 
subtitle: "for Systematic Review"
author: "Charles T. Gray, Matthew Grainger, & Neal Haddaway"
output: 
    ioslides_presentation:
        widescreen: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(sysrevdata)
library(kableExtra)
library(tidyverse)

```

## Interoperable and open | Systematic review {.smaller}

### the problem of too many systematic reviews

88 days after WHO named Covid-19 

88 systematic reviews on Covid-19 [@naumann_too_2021]

### the solution

Cochrane's Covid-19 [living evidence synthesis](https://covid-nma.com/the-project/) [@boutron_covid-nma_2020]  

### practical obstacle

sharing data for future & living evidence synthesis

lack of standardisation of databases

## Human-readable 

```{r}
farming_prod <-
  bufferstrips %>%
  # choose the desired columns
  select(short_title, contains("farmingproductionsystem")) %>%
  # condense variableal
  unite(# name condensed column
    col = "farming_production_system",
    # columns to condense
    contains("farmingproductionsystem"),
    # set separator
    sep = " ||| ",
    # remove NAs
    na.rm = TRUE)

farming_prod %>%
  # count types of farming production system
    count(farming_production_system) %>%
  # order by the count
  arrange(desc(n)) %>%
  head(10) %>% 
    kbl() %>% 
    kable_styling()
    
```

## Machine-readable

```{r}
pivoted_bufferstrips <-
pivot_longer(
  data = bufferstrips,
  cols = -c(1:12, 15, 16),
  names_to = c("category", "subcategory"),
  names_sep = "_",
  values_drop_na = TRUE
) %>%
  select(-value) %>%
  janitor::clean_names()

pivoted_bufferstrips %>%
  select(1,2,3,15, 16)

```


## Toolchains

```{r eval=FALSE,echo=TRUE}
pivoted_bufferstrips <-
pivot_longer(
  data = bufferstrips,
  cols = -c(1:12, 15, 16),
  names_to = c("category", "subcategory"),
  names_sep = "_",
  values_drop_na = TRUE
) %>%
  select(-value) %>%
  janitor::clean_names()

pivoted_bufferstrips %>%
  select(1,2,3,15, 16)
```


## Contributing

- [sysrevdata living resource of toolchains](https://softloud.github.io/sysrevdata/)
- [sysrevdata repository](https://github.com/softloud/sysrevdata)
- [make suggestions, join the discussion: issues](https://github.com/softloud/sysrevdata)

## References
