---
title: "network meta-analysis"
subtitle: "toolchain walkthrough"
author: "Charles T. Gray"
date: "Wednesday 21 October 2020"
bibliography: references.bib 
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, 
    cache = TRUE,
    fig.align = "center"
    )
```

<div class="notes">

objective

- toolchain walkthrough
- learn/bring

</div>

## toolchain walkthrough 

### definition

**toolchain walkthrough** := an _opinionated_ [@parker_opinionated_2017] documentation of a scientific workflow [@gray_codeproofprepare_2019]


### opinionated usage

> _The [tidyverse](https://www.tidyverse.org/) [@wickham_tidyverse_2017] is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures._

### publication

Gray, C. T. (2019). code::proof: Prepare for Most Weather Conditions. Statistics and Data Science, 22–41. https://doi.org/10.1007/978-981-15-1960-4_2

> 

## load packages

```{r packages,message=FALSE}

# toolchain walkthrough
library(multinma) # network meta-analysis
library(nmathresh) # threshold analysis



```

```{r include=FALSE}
conflicted::conflict_prefer("filter", "dplyr")

parkinsons_dat <- parkinsons %>% 
  mutate(
          studyn = str_c("study_", studyn),
          trtn = str_c("trt_", trtn)
  )

# additional packages used in this walkthrough
library(tidyverse) # general purpose
library(kableExtra) # tables
library(patchwork)

```


## example data | Parkinson's disease

```{r eval=FALSE}
# from data documentation
?parkinsons

```


> mean off-time reduction in patients given dopamine agonists as adjunct therapy in Parkinson's disease from 7 trials comparing four active drugs and placebo

```{r}
parkinsons_dat%>%
    # display top values
    head(3)

```



## network meta-analysis model

### meta-analysis

$$
\boldsymbol y \sim \text{normal}(\mu + \boldsymbol \theta, \boldsymbol \sigma^2)\\
\boldsymbol \theta \sim \text{normal}(0, \tau^2)
$$

### bayesian network meta-analysis

\[
\left.
\begin{array}{c r c l}
\text{prior} & \boldsymbol d & \sim & \text{normal}(\boldsymbol d_0, \boldsymbol \Sigma_d)\\
\text{likelihood} & \boldsymbol y | \boldsymbol d & \sim & \text{normal}(\boldsymbol \delta, \boldsymbol V)\\
\text{fixed effects model} &  \boldsymbol \delta &=& \boldsymbol{Xd}
\end{array}
\right\}
\]

## create network object

```{r}
parkinsons_net <-
    set_agd_contrast(
        parkinsons,
        study = studyn,
        trt = trtn,
        y = diff,
        se = se_diff,
        sample_size = n
    )

```



## network meta-analysis model | prior

```{r}
# prior
summary(normal(scale = 100))

```

an example of an area I would like to learn more about; selection of priors. 

## network meta-analysis model | `::nma`

```{r}
# fit model to network object
parkinsons_nma <-
    nma(
        # network object
        parkinsons_net, 
        
        # fixed; for brevity w sensitivity
        trt_effects = "fixed", 
        
        # set prior on treatment contrast with placebo
        prior_trt = normal(scale = 100))

```

# cochrane reporting 

## cochrane reporting | network

```{r}
plot(parkinsons_net, weight_edges = FALSE)

```

## cochrane reporting |  intervention effects

```{r fig.height=3.5}
plot(parkinsons_nma)

```

## cochrane reporting |  intervention effects

```{r fig.height=3.5}
plot_prior_posterior(parkinsons_nma)

```


## cochrane reporting | ranking {.smaller}

```{r fig.width=10, fig.height = 4, fig.align="center"}
posterior_rank_probs(parkinsons_nma) %>% plot() + 
  posterior_rank_probs(parkinsons_nma, cumulative = TRUE) %>% plot()

```

# sensitivity

## leave one out | exclude a random study from the network meta-analysis

```{r include=FALSE}
# choose a random study
study_to_filter <- 
  parkinsons_dat%>% 
  pull(studyn) %>% 
  unique() %>% 
  sample(1)

leave_one_out_plot <-
parkinsons_dat%>% 
  filter(studyn != study_to_filter) %>% 
     set_agd_contrast(
       # data
       .,
        study = studyn,
        trt = trtn,
        y = diff,
        se = se_diff,
        sample_size = n
    ) %>% 
      nma(
        # network object
        .,
        
        # fixed; for brevity w sensitivity
        trt_effects = "fixed", 
        
        # set prior on treatment contrast with placebo
        prior_trt = normal(scale = 100)) %>% 
  plot()


```



```{r echo=FALSE, fig.height=3,fig.align="center"}
leave_one_out_plot

```


```{r}
# excluding study
study_to_filter

```

## leave m out 

```{r combinations, include=FALSE,warning=FALSE}
# get combinations of treatments
# ::combn w ::t might be useful
all_trt_comb <-
  parkinsons_dat%>%
  select(studyn, trtn)

trt_comb <- 
all_trt_comb %>%
  pull(studyn) %>%
  unique() %>%
  map_df(
    .f = function(study) {
      all_trt_comb %>%
        filter(studyn == study) %>%
        pull(trtn) %>%
        combn(2) %>%
        t() %>%
        as_tibble()
    }
  ) %>%
  distinct() %>% 
  filter(V1 == 2)


```

```{r include=FALSE}



# trial function

trial <- function(studies_to_include) {
  
  trial_model <-
  parkinsons_dat%>% 
    filter(studyn %in% studies_to_include) %>% 
     set_agd_contrast(
       # data
       .,
        study = studyn,
        trt = trtn,
        y = diff,
        se = se_diff,
        sample_size = n
    ) %>% 
      nma(
        # network object
        .,
        
        # fixed; for brevity w sensitivity
        trt_effects = "fixed", 
        
        # set prior on treatment contrast with placebo
        prior_trt = normal(scale = 100))
  
  summary(trial_model, par = "d") %>% 
    as_tibble() %>% 
    select(parameter, mean) %>% 
    arrange(mean) %>%
    mutate(rank = row_number(),
           k = length(studies_to_include)) %>% 
    select(-mean)

}

# extract study combinations
studies <-
  parkinsons_dat%>% 
  pull(studyn) %>% 
  unique() 

study_inclusions <-
  3:length(studies) %>% 
  map_df(.f = function(k){
  combn(
    x = studies,
    m = k,
    simplify = FALSE
  ) %>% 
      tibble(
        k = k,
        included_studies = .
      )
  }) 


```

```{r include=FALSE}
rankings <-
  study_inclusions %>%
  pull(included_studies) %>% 
  map_df(trial) 

```

```{r echo=FALSE, fig.height=5, fig.align="center"}
rankings %>%
  # filter(k != 7#, 
  #        # !str_detect(parameter, "4")
  #        ) %>%
  mutate(rank = as_factor(rank)) %>% 
  ggplot(aes(x = parameter, fill = rank)) +
           geom_bar(alpha = 0.6) +
  ggthemes::theme_economist_white() +
  rockthemes::scale_fill_nodoubt() +
  facet_wrap(~k) +
  coord_flip() +
  labs(
    title = "network meta-analysis study subset ranking distributions", 
    subtitle = str_wrap("on treatments for all subsets of 3 or greater of the available 7 studies", 80),
    y = "number of combinations of k studies that measure treatment",
    caption = str_wrap("Bar plots are grouped by k, the number of studies.", 60),
    x = "treatment"
  ) +
  scale_fill_manual(guide = guide_legend(reverse = TRUE),
                    values = palette(rockthemes::nodoubt_pal()))

```



## threshold analysis | arguments

- posterior mean estimates
- likelihood covariance matrix
- posterior covariance matrix
- design matrix of contrasts

## threshold analysis | posterior mean estimates

```{r}
# get posterior mean estimates
post_means <- 
  summary(parkinsons_nma, pars=c("d")) %>% 
  as.data.frame() %>%
  pull("mean")

# print values
post_means

```


## threshold analysis | design matrix {.smaller .columns-2}


```{r height = 5,echo=FALSE}
plot(parkinsons_net, weight_edges = FALSE)

```


```{r echo=FALSE}
#


X <- matrix(data = c(-1, 0, 1, 0,
              -1, 1, 0, 0,
              1, 0, 0, 0,
              0, 0, 1, 0,
              0, 0, 0, 1),
            byrow = TRUE,
            nrow = 5,
            ncol = 4)

X
```


</div>

## threshold | posterior likelihood

```{r}

Sigma_d <- 
  diag(rep(100^2, length(post_means)))



```


## references